Sea: A lightweight data-management library for Big Data scientific computing

### 2022-03-23 paper/sea-comp/paper.tex
updates to sea comp paper

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Hayot-Sasson\MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
\IEEEtitleabstractindextext{%
\begin{abstract}
    The recent influx of scientific computing datasets have shifted scientific
computing from compute intensive to data intensive. Whereas many Big Data
Frameworks exist that minimize the cost of data transfers, few scientific
applications adopt data-management strategies as the effort required to update
the existing applications is significant.

We developed Sea as a means to enable data-management strategies for scientific
applications executing on HPC clusters without needing to update the underlying
applications. Sea leverages GNU C library interception to intercept
POSIX-compliant file system calls made by the applications. We designed a
performance model and evaluated the performance of Sea on a synthetic
data-intensive application processing a representative neuroimaging dataset (the
Big Brain). Our results demonstrate that Sea significantly improves performance,
up to a factor of 3$\times$.

@@

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
\IEEEPARstart{E}{fficient} data-management
strategies have become essential in minimizing the costs of processing Big Data.
Such strategies include data locality and in-memory computing, which minimize
transfers by ensuring compute occurs where the data is located and maintain data
in the fastest available storage (i.e. memory). Big Data frameworks, such as
MapReduce~\cite{dean2008mapreduce}, Apache Spark~\cite{zaharia2016apache} and
Dask~\cite{rocklin2015dask}, provide such data management mechanisms by default.

Until the recent surge in publicly available scientific data, scientific
workflows had been regarded as compute intensive. As a result, standardized
scientific tools typically lack these data-management mechanisms. The sheer
effort required to update the applications to leverage Big Data frameworks (e.g.
changing framework, programming language, training staff) can be a significant
deterrent. Furthermore, Big Data frameworks still require significant work to be
well adapted for the processing of different scientific data, such as imaging
data~\cite{mehta2017comparative}.

@@

Our goal is to provide data-management strategies to scientific tools. We focus
on well-established scientific applications that operate as command-line tools.
These applications are written in a variety of different programming languages
and communicate data between tasks via a POSIX-compliant shared file system.

Scientific tools are typically platform-agnostic with the exception that the
file system must be POSIX-compliant. That is, they can run on local
workstations, high performance computing (HPC) clusters and cloud
infrastructure, with the exception that they cannot communicate with a
cloud-based storage natively. We focus our efforts on HPC clusters as they are
widely used in scientific computing and are currently used for the processing of
scientific Big Data.

Our goal is to provide data-management strategies to scientific tools. We focus
on well-established scientific applications that operate as command-line tools.
These applications are written in a variety of different programming languages
and communicate data between tasks via a POSIX-compliant shared file system.

Scientific tools are typically platform-agnostic with the exception that the
file system must be POSIX-compliant. That is, they can run on local
workstations, high performance computing (HPC) clusters and cloud
infrastructure, with the exception that they cannot communicate with a
cloud-based storage natively. We focus our efforts on HPC clusters as they are
widely used in scientific computing and are currently used for the processing of
scientific Big Data.

@@

Sea\footnote{\url{https://github.com/valhayot/sea}} is an open-source user-space
data-management C++ library for applications running on HPC systems. Its main
aim is to reduce application data transfer costs by leveraging node-local
storage (Figure~\ref{fig:sea-comp:diagram}). Sea redirects files accessed from a user-specified mount point to the
appropriate storage devices using glibc interception. Within an intercepted call
the file will be written to or read from the fastest storage device available.

>>>>

% The paper headers
\markboth{Journal of \LaTeX\ Class Files,~Vol.~14, No.~8, August~2015}%
{Hayot-Sasson\MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
\IEEEtitleabstractindextext{%
\begin{abstract}
    The recent influx of scientific data has lead to scientific
computing transitioning from compute intensive to data intensive. Whereas many Big Data
Frameworks exist that minimize the cost of data transfers, few scientific
applications adopt data-placement strategies as the efforts required to update
the existing applications is significant.

We developed Sea as a means to enable data-placement strategies for scientific
applications executing on HPC clusters without the need to reinstrument workflows.
Sea leverages GNU C library interception to intercept
POSIX-compliant file system calls made by the applications. We designed a
performance model and evaluated the performance of Sea on a synthetic
data-intensive application processing a representative neuroimaging dataset (the
Big Brain). Our results demonstrate that Sea significantly improves performance,
up to a factor of 3$\times$.

@@

\IEEEraisesectionheading{\section{Introduction}\label{sec:introduction}}
\IEEEPARstart{E}{fficient} data-placement
strategies have become essential in minimizing the costs of processing Big Data.
Strategies, such as data locality and in-memory computing, improve application runtime by ensuring
data access occur on the most efficient available storage. Existing Big Data frameworks, such as
MapReduce~\cite{dean2008mapreduce}, Apache Spark~\cite{zaharia2016apache} and
Dask~\cite{rocklin2015dask}, improve runtime performance through the incorporation of these data-placement
strategies.

Until the recent surge in publicly available scientific data, scientific
workflows have been regarded as compute intensive. As a result, standardized
scientific tools typically lack data-placement mechanisms. In contrast, efforts have
been placed on efficient storage representation, such as HDF5~\cite{hdf5} and Zarr~\cite{zarr}, and
the sharing of data, through initiatives like DataLad~\cite{wagner2022fairly}.

Data-placement strategies can complement existing solutions by selecting where data should be accessed at runtime,
facilitating the transfer and processing of large datasets across different infrastructures, and placing the different
file formats at the preferred storage layer to maximize efficiency. However, converting adapting existing scientific frameworks
and applications to Big Data frameworks requires significant effort, necessitating the rewrite of many well established applications.
Furthermore, as noted in \cite{mehta2017comparative}, Big Data frameworks are still difficult to use for the processing of certain
scientific data, such as imaging data.

@@

Our goal is to provide a data-placement add-on to standardized scientific computing applications executing on High Performance Computing (HPC)
clusters. The tools of particular interest to us consist of various tasks that communicate to each other via a POSIX-compliant file system. Whereas
they can be platform-agnostic, users typically leverage them within an HPC context, particularly when processing large volumes of data, due to their
inherent reliance on POSIX-compliance.

We developed Sea, a user-space data-placement library, that provides efficient data placement without reinstrumentation of the underlying application.
Given a predefined storage hierarchy, Sea can prefetch, cache, flush and evict data to and from a permanent storage location. To evaluate Sea,
we have designed a performance model, as well as evaluate its performance on a synthetic Big Data application running on a representative neuroimaging dataset.

@@

Sea\footnote{\url{https://github.com/valhayot/sea}} is an open-source user-space
data-placement C++ library for applications running on HPC systems. Its main
aim is to reduce application data transfer costs by leveraging node-local
storage (Figure~\ref{fig:sea-comp:diagram}). Sea redirects files accessed from a user-specified mount point to the
appropriate storage devices using glibc interception. Within an intercepted call
the file will be written to or read from the fastest storage device available.
